---
title: "Assignment_6_Group_3"
author: "Jin Run, Jiang Tao, Mu Xin, Dang Vu Nguyen and Gunn Wei Lim"
date: "9 March 2019"
output: html_document
---
ABC Airlines has 100 economic class seats available in its flight from Singapore to Shanghai.
Full price fare is 590 dollar. ABC is considering selling some seats at a discounted price of 450 dollar.
The demand of the flight follows normal distribution with mean 110 and standard deviation 30.
Based on past experience, ABC is able to sell all discounted tickets as long as there is demand.
However, only 60% of the customers will buy full fare ticket when no more discounted tickets available.

1) Build a simulation model to decide the optimal number of seats to be reserved as full fare seats.

```{r}

iteration <- function(full.fare.seats,iters,cap)
{
  #iterate demand for flight seats
  seats <- trunc(rnorm(iters,110,30))

  #Number of discount seats, if less than cap-full.fare.seats then discounts seats = seats.filled.
  discount.seats <- sapply(seats,function(x) min(cap-full.fare.seats,x))
  
  #If number of seats < maximum discounts seats, full fare seats = 0.  if number of demand for discount seats 
  #& demand for full fare seats <capacity, then it's 60% of seats minus max discounts seats. Else, will serve
  #maximum full fare seats

  filled.full.fare.seats <- ifelse(seats<(cap-full.fare.seats),0,ifelse(0.6*seats+0.4*(cap-full.fare.seats)<cap,trunc(0.6*(seats-cap+full.fare.seats)),full.fare.seats))
  seatdf <- data.frame(full.fare=filled.full.fare.seats,discount=discount.seats)
  seatdf$revenue <- 590*seatdf$full.fare + 450*seatdf$discount
  return(mean(seatdf$revenue))
}

simulation1 <- function(iters, cap){
  library(ggplot2)
  results <- sapply(1:cap,iteration,iters,cap)
  df <- data.frame(number.of.full.fare.seats=1:cap,average.revenue=results)
  optimalSeat = which.max(results)
  maxRev=max(results)
  print(optimalSeat)
  ggplot(df, aes(x=number.of.full.fare.seats,y=average.revenue)) + geom_line(group = 'average.revenue') + geom_vline(xintercept = optimalSeat,color='red',linetype='dashed') + geom_text(aes(x=optimalSeat-3, label=paste0("Optimal at ",optimalSeat,"with revenue of",maxRev), y=maxRev/2), colour="blue")
}

simulation1(2000000,100)

```

2) What if the mean demand change? How will it affect the full fare seats allocation?

```{r}
iteration2 <- function(full.fare.seats,mean.demand,iters,cap)
{
  #iterate demand for flight seats
  seats <- trunc(rnorm(iters,mean.demand,30))

  #Number of discount seats, if less than cap-full.fare.seats then discounts seats = seats.filled.
  discount.seats <- sapply(seats,function(x) min(cap-full.fare.seats,x))
  
  #If number of seats < maximum discounts seats, full fare seats = 0.  if number of demand for discount seats 
  #& demand for full fare seats <capacity, then it's 60% of seats minus max discounts seats. Else, will serve
  #maximum full fare seats

  filled.full.fare.seats <- ifelse(seats<(cap-full.fare.seats),0,ifelse(0.6*seats+0.4*(cap-full.fare.seats)<cap,trunc(0.6*(seats-cap+full.fare.seats)),full.fare.seats))
  seatdf <- data.frame(full.fare=filled.full.fare.seats,discount=discount.seats)
  seatdf$revenue <- 590*seatdf$full.fare + 450*seatdf$discount
  return(mean(seatdf$revenue))
}


#Finding optimal number of full fare seats given mean demand
simulation2 <- function(mean.demand,iters, cap){
  results <- sapply(1:cap,iteration2,mean.demand,iters,cap)
  optimalSeat = which.max(results)
  return(optimalSeat)
}


#to Plot the correlation between optimal number of full fare seats for each mean demand
demand.simu <- function(iters,cap){
  library(ggplot2)
  results <- sapply(50:200,simulation2,iters,cap)
  df <- data.frame(seats.demand=50:200,optimal.full.fare.seats=results)
  ggplot(df, aes(y=optimal.full.fare.seats,x=seats.demand)) + geom_line(group = 'seats.demand')
}

demand.simu(10000,100)

```

3) What if discount percentage change? How will it affect the full fare seats allocation?

```{r}
iteration3 <- function(full.fare.seats,discount.percent,iters,cap)
{
  #iterate demand for flight seats
  seats <- trunc(rnorm(iters,110,30))

  #Number of discount seats, if less than cap-full.fare.seats then discounts seats = seats.filled.
  discount.seats <- sapply(seats,function(x) min(cap-full.fare.seats,x))
  
  #If number of seats < maximum discounts seats, full fare seats = 0.  if number of demand for discount seats 
  #& demand for full fare seats <capacity, then it's 60% of seats minus max discounts seats. Else, will serve
  #maximum full fare seats

  filled.full.fare.seats <- ifelse(seats<(cap-full.fare.seats),0,ifelse(0.6*seats+0.4*(cap-full.fare.seats)<cap,trunc(0.6*(seats-cap+full.fare.seats)),full.fare.seats))
  seatdf <- data.frame(full.fare=filled.full.fare.seats,discount=discount.seats)
  #calculating average revenue
  seatdf$revenue <- 590*seatdf$full.fare + 590*discount.percent*seatdf$discount
  return(mean(seatdf$revenue))
}
iteration3(30,0.1,100,100)

#Finding optimal number of full fare seats given discount percentage
simulation3 <- function(discount.percent,iters, cap){
  results <- sapply(1:cap,iteration3,discount.percent,iters,cap)
  optimalSeat <- which.max(results)
  return(optimalSeat)
}
simulation3(0.1,10,100)

#to Plot the correlation between optimal number of full fare seats for each discount percentage
discount.simu <- function(iters,cap){
  library(ggplot2)
  results <- sapply(seq(0.3, 1, by = 0.05),simulation3,iters,cap)
  df <- data.frame(discount.percentage=seq(0.3, 1, by = 0.05),optimal.full.fare.seats=results)
  ggplot(df, aes(x=discount.percentage,y=optimal.full.fare.seats)) + geom_line(group = 'optimal.full.fare.seats')
}

discount.simu(10000,100)
```

4) What is the best combination of seat allocation and discount percentage with respect to a particular demand?
```{r}
knitr::opts_chunk$set(echo=TRUE,warning=FALSE,message = FALSE)
library(ggplot2)
library(dplyr)
library(reshape2)
GetSingleRev <- function(reserved,seats,demand,disc.p,full.p){
  discount.seats <- seats - reserved
  if (demand < discount.seats){
    rev <- demand * disc.p
  }else{ 
    if(0.6*(demand-discount.seats) <= reserved ){
      rev <- discount.seats * disc.p + round(0.6*(demand-discount.seats))*full.p
    }
    else{
      rev <- discount.seats * disc.p + reserved*full.p
    }
  }
  return (rev)
}

#Input: demand,disc.p, 
#Iterate: reserved seats 
#Output: revenues for all posible reserved seats 
GetRevForAllReserved <-function(seats,demand,disc.p,full.p){
  revenues <- sapply(0:seats,GetSingleRev,seats,demand,disc.p,full.p)
  AllReserved <- data.frame(Reserved= 0:seats, revenues)
  return(AllReserved)
}

#Input: disc.p
#Iterate: demands
#Outout: revenues for all mean posibilities 
GetRevForAllDemand <-function(seats, min.mean,max.mean,sd,disc.p,full.p){
  outcome <- data.frame()
  for (i in seq(min.mean,max.mean, by=5)){
    demand <- rnorm(1,i,sd)
    demand <-ifelse(demand<0,0,demand)
    #One line of result is a list of revenues for all Reserved numbers 
    result <- GetRevForAllReserved(seats,demand,disc.p,full.p)
    current.outcome <- cbind(result,i)
    outcome <- rbind(outcome,current.outcome)
  }
  names(outcome) <-c("Reserved","Revenue","Demand Mean")
  return(outcome)
}

#Fix demand,iterate for each percentage, find best # of reserved seats 
RevenueForEachPercentage <-function(seats, min.percent, max.percent, min.mean, max.mean, sd, full.p){
  outcome <- data.frame()
  for (i in seq(min.percent,max.percent,by=5)){
    disc.p <- (i)/100*full.p
    #Each Result is a data frame 
    result <- GetRevForAllDemand(seats, min.mean, max.mean,sd,disc.p,full.p)
    current.outcome <- cbind(result,i)
    outcome <- rbind(outcome,current.outcome)
  }
  names(outcome) <-c("Reserved","Revenue","Demand Mean","Discount Percentage")
  return(outcome)
}

#For each mean, what is the best reserved seat 
simulation <- function(iter, seats, min.percent,max.percent,min.mean,max.mean,sd,full.p) {
  AllIterResult <- data.frame()
  for (i in 1:iter) {
    OneIterResult <-
      RevenueForEachPercentage(seats, min.percent, max.percent, min.mean, max.mean, sd, full.p)
      #names(OneIterResult) <- c("Percent", "disc.p", "revenue")
      AllIterResult <- rbind(AllIterResult, OneIterResult)
  }
  #names(AllIterResult) <- c("Reserved","Revenue","Demand Mean","Discount Percentage")
  
  #Take average of all revenues from all iterations 
  SumAllIterResult <- with(AllIterResult,aggregate(AllIterResult$Revenue, list(AllIterResult$Reserved,AllIterResult$`Demand Mean`,AllIterResult$`Discount Percentage`), mean))
  names(SumAllIterResult) <- c("Reserved","Demand Mean","Discount Percentage","Revenue")
  #Find the optimal reservation for each demand and discount percentage
  Optimal <- SumAllIterResult %>% 
    group_by(`Demand Mean`,`Discount Percentage`) %>% 
    filter(Revenue==max(Revenue))
  Optimal <- Optimal[,-4]
  
  resultMatrix <-round(dcast(Optimal,`Demand Mean` ~ `Discount Percentage`,value.var="Reserved",fun.aggregate = mean))
  
  p <- ggplot(Optimal, aes(x = `Demand Mean`, y = `Discount Percentage`)) +
  geom_raster(aes(fill=Reserved)) +
  scale_fill_gradient(low="red", high="blue") +
  labs(x="Demand Mean", y="Discount Percentage", title="Matrix showing optimal number of seats reserved for full price") +
  theme_bw() + theme(axis.text.x=element_text(size=9, angle=0, vjust=0.3),
                     axis.text.y=element_text(size=9),
                     plot.title=element_text(size=11))
  print(resultMatrix)
  print(p)
  }
simulation(100,100,0,100,50,200,30,590)
```